Depletion capabilities in the OpenMC Monte Carlo particle transport code

### 2020-06-24 manuscript.tex
More proofreading edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[charter]{mathdesign} % Type-1 Latin Modern modern
\usepackage[T1]{fontenc}         % Use T1 encoding instead of OT1
\usepackage[utf8]{inputenc}      % Use UTF8 input encoding
\usepackage{microtype}           % Improve typography
\usepackage{booktabs}            % Publication quality tables
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{multirow}
\usepackage{upgreek}

@@

To validate the depletion solver in OpenMC, we have carried out simulations
using the \texttt{openmc.deplete} module as well as a developmental version of
Serpent 2.1.32 on two simple problems. The first problem is a pressurized water
reactor (PWR) pincell and the second problem is a single assembly of a
sodium-cooled fast reactor (SFR). These problems were selected to cover both
thermal and fast neutron spectra. While we had originally attempted to use
Serpent 2.1.31, early comparisons between OpenMC and this version of Serpent
revealed significant differences in the predicted concentration of $^{97}$Mo.
This was found to be due to $^{97\text{m}}$Y having a non-zero independent yield
and a zero cumulative yield in the FPY sublibrary file for $^{235}$U fission in
ENDF/B-VII.1. The zero cumulative yield caused Serpent 2.1.31 to ignore it
altogether, which resulted in significantly less production of $^{97}$Mo. The
developmental version of Serpent 2.1.32 no longer ignores this particular
fission product despite the erroneous cumulative yield.

@@

Given the remarks above, it is of utmost importance to ensure that the same
nuclear data are being used in a cross-code comparison. For a depletion
simulation, we must ensure that the incident neutron cross sections and
secondary angle/energy distributions, radioactive decay data, and fission
product yields are the same for both codes. In the simulations performed, all
data was based on the ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}.
Incident neutron cross sections from the ENDF/B-VII.1 library processed by the
LANL nuclear data team~\citep{conlin2013lanl}, known as endf71x, were used for
transport simulations in both OpenMC and Serpent. Serpent is able to use the ACE
files from the endf71x library directly while OpenMC uses an HDF5 format that
requires conversion of the ACE files first. For the PWR pincell problem,
$S(\alpha,\beta)$ data is also needed for hydrogen in light water; because
Serpent does not yet handle the continuous representation used in the ENDF71SaB
library, the light water data from ENDF/B-VII.0 was used instead. We note that
the light water evaluation didn't actually change between ENDF/B-VII.0 and
ENDF/B-VII.1.

@@

To ensure that OpenMC and Serpent perform power normalization in the same
manner, the default fission heating values from Serpent were used in OpenMC. By
default, the energy release per fission in $^{235}U$ in Serpent is set to
\SI{202.27}{\mega\electronvolt}, and the values for other actinides is scaled
based on the Q value for fission listed in MF=3 in the ENDF evaluation. Without
the same heating values, the power normalization might result in OpenMC
depleting materials too quickly or too slowly.

Serpent uses energy-independent capture branching ratios by default that are
based on the JEFF 3.1 activation file using a PWR flux spectrum. For the PWR
problem, these energy-independent branching ratios were used in OpenMC. For the
SFR problem, Serpent was used to first generate representative capture branching
ratios in the fast flux spectrum using Serpent's energy-dependent capture
branching ratio treatment. Then, the energy-independent branching ratios were
set in both Serpent and OpenMC. As discussed in \cref{sec:fpy}, OpenMC also has
several methods for treating the energy dependence of FPYs. The OpenMC
simulations were carried out with the three different methods to see what effect
they have on resulting isotopic compositions as compared with Serpent.

@@

The PWR pincell problem is a single 2.4\% enriched PWR pincell from the BEAVRS
benchmark~\citep{horelik2013mc,horelik2018mit}. Material compositions and
dimensions from the BEAVRS benchmark were used without modification. The pincell
extends to positive/negative infinity in the $z$ direction and has reflective
boundary conditions in the $x$ and $y$ directions. A linear heating rate of
\SI{174}{\watt\per\centi\meter} was used for power normalization. The pin was
depleted to 50 MWd/kg using the following burnup schedule:
\begin{itemize}
  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps;
  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg;
  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg.
\end{itemize}
The short timesteps at the beginning are used to accurately capture the buildup
of xenon isotopes. In both codes, the predictor method was used for time
integration. While the predictor method is the least accurate time integration
method, the purpose of the comparison is to confirm that OpenMC and Serpent
produce comparable results, irrespective of the absolute values.

Five coupled transport--depletion simulations were carried out with OpenMC to test
a variety of combinations of depletion chains and methods for handling FPY
energy dependence:
\begin{enumerate}
  \item Full depletion chain, thermal FPYs
  \item Full depletion chain, interpolated FPYs
  \item Reduced depletion chain, thermal FPYs
  \item CASL depletion chain, thermal FPYs
  %\item CASL depletion chain, FPYs based on cutoff method
  \item CASL depletion chain, interpolated FPYs
\end{enumerate}
Because Serpent doesn't have an option for changing the depletion chain or the
method for handling FPY energy dependence, a single coupled transport--depletion
simulation with Serpent was carried out. For all simulations, 20 inactive
batches, 200 active batches, and \num{4.0e6} particles per batch were used,
which resulted in an uncertainty on the effective multiplication factor of about
3--4 pcm.

\Cref{fig:pwr-keff} shows the difference between $k_\text{eff}$ calculated by
OpenMC and Serpent. The blue points correspond to results using the full
depletion chain, the red points correspond to points using the reduced chain,
and the green points correspond to results using the CASL chain. For each chain,
the use of interpolated FPYs is denoted by circles and thermal FPYs by x's. At
zero days, the difference in $k_\text{eff}$ between OpenMC and Serpent is not
statistically significant. As the burnup increases, the difference for the full
chain cases rises to about 20 pcm. The reduced chain case exhibits the exact
same behavior as the full chain as one would expect. Use of the CASL chain
results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
about 40 pcm. We also see that for both the full and CASL chains, changing the
method for handling FPY energy dependence doesn't appear to have much of an
affect on $k_\text{eff}$, which is primarily driven by the prediction of
actinide concentrations.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_keff.pdf}
  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
  between OpenMC and Serpent for the PWR pincell model. While not shown, the one
  standard deviation uncertainty in the difference is about 3.5 pcm.}
  \label{fig:pwr-keff}
\end{figure}

@@

\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concetrations
when using the full depletion chain along with thermal FPYs. In this case,
although most fission products are within 1\% of Serpent, they are
systematically overpredicted other than a few nuclides. Because Serpent uses the
average fission energy to interpolate between two sets of FPYs, the systematic
difference is actually expected when using thermal FPYs in OpenMC. As the
incident energy of a neutron causing fission increases, the ``saddle'' and
``wings'' in the FPY curve tend to increase significantly, and the peaks around
$A=95$ and $A=135$ tend to slightly decrease. This corresponds to exactly what
is observed in \cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in
OpenMC for this case, we expect that fission products near one of the two peaks
of the FPY curve will be slightly overpredicted relative to Serpent, and fission
products away from the peak will be underpredicted. This can be confirmed for
particular nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain,
the nuclides with the highest independent yields are $^{109}$Tc and $^{109}$Mo.
Both these nuclides have independent yields that increase significantly with
increasing neutron energy. The same situation occurs for $^{129}$I, whose
predominant precursor is $^{129}$Sn. The independent yields for these three
precursor nuclides are shown in \cref{tab:fp-yields}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_thermal.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  full depletion chain and thermal FPYs.}
  \label{fig:pwr-fp-full-thermal}
\end{figure}
\begin{table}[htbp]
  \centering
  \caption{Independent yields of select fission products due to fission in
  $^{235}$U (data from the ENDF/B-VII.1 FPY sublibrary).}
  \label{tab:fp-yields}
  \begin{tabular}{cccc}
    \toprule
    \textbf{Energy Spectrum} & \textbf{$^{109}$Tc} & \textbf{$^{109}$Mo} & \textbf{$^{129}$Sn} \\
    \midrule
    Thermal & 0.00013 & 0.00015 & 0.00230 \\
    Fast & 0.00034 & 0.00042 & 0.00550 \\
    Fusion & 0.00598 & 0.00109 & 0.01162 \\
    \bottomrule
  \end{tabular}
\end{table}

\Cref{fig:pwr-fp-casl-average} shows the relative difference in predicted
fission product concentrations between OpenMC and Serpent when using the CASL
depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
almost all fission products are predicted to well within 1\% at 50 MWd/kg.
However, in this case there is one clear outlier: $^{85}$Kr. Upon further
investigation, it was determined that the root cause for this difference is the
fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
the cumulative yield of $^{85}$Kr is known to be inconsistent with the
independent yields of the $A=85$ mass chain~\citep{pigni2015nds}. Serpent relies
solely on the independent yield in its depletion solver (as does the full
depletion chain). Therefore, there is nothing inherently wrong with the CASL
depletion chain or OpenMC; it is simply a matter of the underlying nuclear data
being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
depletion chain shown in \cref{fig:pwr-keff}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  CASL depletion chain and interpolated FPYs.}
  \label{fig:pwr-fp-casl-average}
\end{figure}

@@

Unlike the PWR pincell model, which only contains materials at a single
temperature, the SFR assembly model uses a higher temperature in the fuel than
in other materials. In OpenMC, the temperature-dependence of neutron cross
sections was handled by interpolating data between neighboring temperatures.
Serpent uses an explicit temperature treatment known as target motion
sampling~\cite{viitanen2012nse}. However, we note that for this problem, the
temperature dependence of unresolved resonance probability tables becomes
important because of the fast flux spectrum, but the temperature treatment in
Serpent 2.1.31 does not handle probability tables.

Four coupled transport--depletion simulations were carried out with OpenMC for
the SFR assembly model:
\begin{enumerate}
  \item Full depletion chain, interpolated FPYs, probability tables on
  \item CASL depletion chain, interpolated FPYs, probability tables on
  \item Full depletion chain, interpolated FPYs, probability tables off
  \item CASL depletion chain, interpolated FPYs, probability tables off
\end{enumerate}
The results for the PWR pincell model indicated that the reduced depletion chain
results in identical behavior to the full depletion chain, so we have not
included a case for the SFR assembly using the reduced depletion chain. Two
coupled transport--depletion simulation with Serpent was carried out, one with
probabilty tables turned on and one with them turned off. For all simulations,
20 inactive batches, 200 active batches, and \num{4.0e6} particles per batch
were used, which resulted in an uncertainty on the effective multiplication
factor of about 3--4 pcm.

@@

\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ over time
calculated by OpenMC and Serpent. The blue points correspond to results using
the full depletion chain, and the green points correspond to results using the
CASL chain. For each chain, simulations with probability tables on are denoted
by circles and simulations without probability tables are denoted by squares.
Unlike in the PWR pincell case, there is a statistically significant difference
in $k_\text{eff}$ at zero days. Let us first consider the case with probability
tables turned off. In this case, $k_\text{eff}$ calculated by OpenMC is about 20
pcm higher than that of Serpent. When using the full depletion chain, there is
no change in the difference between OpenMC and Serpent. When using the CASL
depletion chain, the difference increases by about 20--30 pcm, similar to the
behavior observed in \cref{fig:pwr-keff}. With probability tables turned on,
however, the situation is quite different. At zero days, $k_\text{eff}$
calculated by OpenMC is 20 pcm lower than the value calculated by Serpent. As
the burnup increases, there is an increasing positive bias between OpenMC and
Serpent. To understand the cause of this difference, it is instructive to look
at the differences in isotopic compositions.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/sfr_keff.pdf}
  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
  between OpenMC and Serpent for the SFR assembly model. While not shown, the one
  standard deviation uncertainty in the difference is about 3.5 pcm.}
  \label{fig:sfr-keff}
\end{figure}

@@

With probability tables turned on, the differences between OpenMC and Serpent in
fission product concentrations are mostly the same as when probability tables
were turned off, as shown in \cref{fig:sfr-fp-nop}. We note that because the
difference in $^{238}$U and $^{239}$Pu was still less than 0.2\% in
\cref{fig:sfr-actinides}, it is expected that the use of probability tables
doesn't have a large effect on the fission product concentrations. Again, most
fission product concentrations agree to within 1\%, with $^{109}$Ag and
$^{85}$Kr being the notable outliers.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average_nop.pdf}
    \caption{Full depletion chain}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average_nop.pdf}
    \caption{CASL depletion chain}
  \end{subfigure}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables off.}
  \label{fig:sfr-fp-nop}
\end{figure}

@@

A new depletion solver has been implemented in the OpenMC Monte Carlo particle
transport code. The depletion solver is implemented primarily in Python 3 and
interfaces with OpenMC's transport solver through Python bindings to an
underlying C++ API, which enables data transfers to happen in memory rather than
through filesystem I/O. The design of the depletion solver relies on a set of
abstract classes that provide modularity and separation between difference
components, including the time integration methods, the matrix exponential
solver, and the transport operator. In principle, this modularity would allow
the depletion system to be used with another transport code. Rather than relying
on a single time integration method, OpenMC implements several time integration
schemes including the predictor method, standard predictor--corrector methods,
stochastic implicit Euler integrators, and several novel
methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
OpenMC relies on the incomplete partial fraction form of CRAM by default.

@@

We found that when comparing results on the SFR assembly model, larger
differences in $k_\text{eff}$ and isotopic concentrations were observed when
unresolved resonance region probability tables were turned on. These differences
were found to be related to how the temperature dependence of cross sections is
treated in the unresolved resonance region. In OpenMC, when the temperature of a
material is in between two temperatures at which probability tables are
tabulated, statistical interpolation between the tables is performed. Serpent,
on the other hand, uses probability tables at the lower tabulated temperature.
For the SFR model, this results in lower parasitic neutron absorption in
$^{238}$U, lower production of $^{239}$Pu, and ultimately a lower
$k_\text{eff}$. After depleting for 360 days, there was a 100 pcm difference
between OpenMC and Serpent due to this effect. Although this is not a huge
difference, users who are simulating problems with a non-uniform temperature
distribution should be aware of this behavior.

@@

Discuss limitations (not validated against experiment, only as good as data).
Improvements in runtime are needed. Only tested for fission applications.
Further development and testing needed for other application areas, such as
transmutation in fusion reactors.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[charter]{mathdesign} % Type-1 Latin Modern modern
\usepackage[T1]{fontenc}         % Use T1 encoding instead of OT1
\usepackage[utf8]{inputenc}      % Use UTF8 input encoding
\usepackage{microtype}           % Improve typography
\usepackage{booktabs}            % Publication quality tables
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[per-mode=symbol]{siunitx}
\usepackage{multirow}
\usepackage{upgreek}

@@

To validate the depletion solver in OpenMC, we have carried out simulations
using the \texttt{openmc.deplete} module as well as a developmental version of
Serpent 2.1.32 on two simple problems. The first problem is a pressurized water
reactor (PWR) pincell and the second problem is a single assembly of a
sodium-cooled fast reactor (SFR). These problems were selected to cover both
thermal and fast neutron spectra. While we had originally attempted to use
Serpent 2.1.31, preliminary comparisons between OpenMC and this version of
Serpent revealed significant differences in the predicted concentration of
$^{97}$Mo. This difference was found to be due to $^{97\text{m}}$Y having a
non-zero independent yield and a zero cumulative yield in the FPY sublibrary
file for $^{235}$U fission in ENDF/B-VII.1. The zero cumulative yield caused
Serpent 2.1.31 to ignore it altogether, which resulted in significantly less
production of $^{97}$Mo. The developmental version of Serpent 2.1.32 no longer
ignores this particular fission product despite the erroneous cumulative yield.

@@

Given the remarks above, it is of utmost importance to ensure that the same
nuclear data are being used in a cross-code comparison. For a depletion
simulation, we must ensure that the incident neutron cross sections and
secondary angle/energy distributions, radioactive decay data, and FPYs are the
same for both codes. In the simulations performed, all data was based on the
ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}. Incident neutron
cross sections from the ENDF/B-VII.1 library processed by the LANL nuclear data
team~\citep{conlin2013lanl}, known as endf71x, were used for transport
simulations in both OpenMC and Serpent. Serpent is able to use the ACE files
from the endf71x library directly while OpenMC uses an HDF5 format that requires
conversion of the ACE files first. For the PWR pincell problem,
$S(\alpha,\beta)$ thermal scattering data is also needed for hydrogen in light
water; because Serpent does not yet handle the continuous representation used in
the ENDF71SaB library, the light water data from ENDF/B-VII.0 was used instead.
We note that the light water evaluation didn't actually change between
ENDF/B-VII.0 and ENDF/B-VII.1.

@@

To ensure that OpenMC and Serpent perform power normalization in the same
manner, the default fission heating values from Serpent were used in OpenMC. By
default, the energy release per fission in $^{235}$U in Serpent is set to
\SI{202.27}{\mega\electronvolt}, and the values for other actinides are scaled
based on the Q value for fission listed in MF=3 in the ENDF evaluation. Without
the same heating values, the power normalization might result in OpenMC
depleting materials too quickly or too slowly relative to Serpent.

Serpent uses energy-independent capture branching ratios by default that are
based on the JEFF 3.1 activation file using a PWR flux spectrum. For the PWR
problem, these energy-independent branching ratios were used in OpenMC. For the
SFR problem, Serpent was used to first generate representative capture branching
ratios in the fast flux spectrum using Serpent's energy-dependent capture
branching ratio treatment. Then, the energy-independent branching ratios were
set in both Serpent and OpenMC.

@@

The PWR pincell problem is a single 2.4\% enriched PWR pincell from the BEAVRS
benchmark~\citep{horelik2013mc,horelik2018mit}. Material compositions and
dimensions from the BEAVRS benchmark were used without modification. The pincell
extends to positive/negative infinity in the $z$ direction and has reflective
boundary conditions in the $x$ and $y$ directions. A linear heating rate of
\SI{174}{\watt\per\centi\meter} was used for power normalization. The pin was
depleted to 50 MWd/kg using the following burnup schedule:
\begin{itemize}
  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps;
  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg;
  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg.
\end{itemize}
The short timesteps at the beginning are used to accurately capture the buildup
of xenon isotopes. In both codes, the predictor method was used for time
integration. While the predictor method is the least accurate time integration
method, the purpose of the comparison is to confirm that OpenMC and Serpent
produce comparable results irrespective of the absolute values.

Five coupled transport--depletion simulations were carried out with OpenMC to test
a variety of combinations of depletion chains and methods for handling FPY
energy dependence:
\begin{enumerate}
  \item Full depletion chain, thermal FPYs
  \item Full depletion chain, interpolated FPYs
  \item Reduced depletion chain, interpolated FPYs
  \item CASL depletion chain, thermal FPYs
  %\item CASL depletion chain, FPYs based on cutoff method
  \item CASL depletion chain, interpolated FPYs
\end{enumerate}
Because Serpent doesn't have an option for changing the depletion chain or the
method for handling FPY energy dependence, a single coupled transport--depletion
simulation with Serpent was carried out. For all simulations, 20 inactive
batches, 200 active batches, and \num{4.0e6} particles per batch were used,
which resulted in an uncertainty on the effective multiplication factor of about
3--4 pcm.

\Cref{fig:pwr-keff} shows the difference between $k_\text{eff}$ calculated by
OpenMC and Serpent. The blue points correspond to results using the full
depletion chain, the red points correspond to points using the reduced chain,
and the green points correspond to results using the CASL chain. For each chain,
the use of interpolated FPYs is denoted by circles and thermal FPYs by x's. At
zero days, the difference in $k_\text{eff}$ between OpenMC and Serpent is not
statistically significant. As the burnup increases, the difference for the full
chain cases rises to about 20 pcm. The reduced chain case exhibits the exact
same behavior as the full chain as one would expect. Use of the CASL chain
results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
about 40 pcm. We also see that for both the full and CASL chains, changing the
method for handling FPY energy dependence doesn't appear to have much of an
effect on $k_\text{eff}$, which is primarily driven by the prediction of
actinide concentrations.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_keff.pdf}
  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
  between OpenMC and Serpent for the PWR pincell model. While not shown, the one
  standard deviation uncertainty in the difference is about 3.5 pcm.}
  \label{fig:pwr-keff}
\end{figure}

@@

\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concetrations
when using the full depletion chain along with thermal FPYs. In this case,
although most fission products are within 1\% of Serpent, they are
systematically overpredicted other than a few nuclides. Because Serpent uses the
average fission energy to interpolate between two sets of FPYs, the systematic
difference is expected when using thermal FPYs in OpenMC. As the incident energy
of a neutron causing fission increases, the ``saddle'' and ``wings'' in the FPY
curve tend to increase significantly, and the peaks around $A=95$ and $A=135$
tend to slightly decrease. This corresponds to exactly what is observed in
\cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in OpenMC for
this case, we expect that fission products near one of the two peaks of the FPY
curve will be slightly overpredicted relative to Serpent, and fission products
away from the peak will be underpredicted. This can be confirmed for particular
nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain, the nuclides
with the highest independent yields are $^{109}$Tc and $^{109}$Mo. Both these
nuclides have independent yields that increase significantly with increasing
neutron energy. The same situation occurs for $^{129}$I, whose predominant
precursor is $^{129}$Sn. The independent yields for these three precursor
nuclides are shown in \cref{tab:fp-yields}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_thermal.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  full depletion chain and thermal FPYs.}
  \label{fig:pwr-fp-full-thermal}
\end{figure}
\begin{table}[htbp]
  \centering
  \caption{Independent yields of select fission products due to fission in
  $^{235}$U (data from the ENDF/B-VII.1 FPY sublibrary).}
  \label{tab:fp-yields}
  \begin{tabular}{cccc}
    \toprule
    \textbf{Energy Spectrum} & \textbf{$^{109}$Tc} & \textbf{$^{109}$Mo} & \textbf{$^{129}$Sn} \\
    \midrule
    Thermal & 0.00013 & 0.00015 & 0.00230 \\
    Fast & 0.00034 & 0.00042 & 0.00550 \\
    Fusion & 0.00598 & 0.00109 & 0.01162 \\
    \bottomrule
  \end{tabular}
\end{table}

\Cref{fig:pwr-fp-casl-average} shows the relative difference in predicted
fission product concentrations between OpenMC and Serpent when using the CASL
depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
almost all fission products are predicted to well within 1\% at 50 MWd/kg.
However, in this case there is one clear outlier: $^{85}$Kr. Upon further
investigation, it was determined that the root cause for this difference is the
fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
the cumulative yield of $^{85}$Kr is known to be inconsistent with the
independent yields of the $A=85$ mass chain in
ENDF/B-VII.1~\citep{pigni2015nds}. Serpent relies solely on independent yields
in its depletion solver (as does the full depletion chain). We can conclude,
then, that there is nothing inherently wrong with the CASL depletion chain or
OpenMC; it is simply a matter of the underlying nuclear data being inconsistent.
The 5\% underprediction of $^{85}$Kr is also likely responsible for the slightly
higher difference in $k_\text{eff}$ for the CASL depletion chain shown in
\cref{fig:pwr-keff}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  CASL depletion chain and interpolated FPYs.}
  \label{fig:pwr-fp-casl-average}
\end{figure}

@@

Unlike the PWR pincell model, which only contains materials at a single
temperature, the SFR assembly model uses a higher temperature in the fuel than
in other materials. In OpenMC, the temperature-dependence of neutron cross
sections was handled by interpolating data between neighboring temperatures.
Serpent uses an explicit temperature treatment known as target motion
sampling~\citep{viitanen2012nse}. However, we note that for this problem, the
temperature dependence of unresolved resonance probability tables becomes
important because of the fast flux spectrum, but the temperature treatment in
Serpent 2.1.31 does not handle probability tables.

Four coupled transport--depletion simulations were carried out with OpenMC for
the SFR assembly model:
\begin{enumerate}
  \item Full depletion chain, interpolated FPYs, probability tables on
  \item CASL depletion chain, interpolated FPYs, probability tables on
  \item Full depletion chain, interpolated FPYs, probability tables off
  \item CASL depletion chain, interpolated FPYs, probability tables off
\end{enumerate}
The results for the PWR pincell model indicated that the reduced depletion chain
results in identical behavior to the full depletion chain, so we have not
included a case for the SFR assembly using the reduced depletion chain. Two
coupled transport--depletion simulation with Serpent was carried out, one with
probabilty tables turned on and one with them turned off. For all simulations,
20 inactive batches, 200 active batches, and \num{4.0e6} particles per batch
were used, which resulted in an uncertainty on the effective multiplication
factor of 3--4 pcm.

@@

\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ calculated by
OpenMC and Serpent as the assembly is depleted. The blue points correspond to
results using the full depletion chain, and the green points correspond to
results using the CASL chain. For each chain, simulations with probability
tables on are denoted by circles and simulations without probability tables are
denoted by squares. Unlike in the PWR pincell case, there is a statistically
significant difference in $k_\text{eff}$ at zero days. Let us first consider the
case with probability tables turned off. In this case, $k_\text{eff}$ calculated
by OpenMC is about 20 pcm higher than that of Serpent. When using the full
depletion chain, there is no change in the difference between OpenMC and Serpent
with increasing burnup. When using the CASL depletion chain, the difference
increases by about 20--30 pcm, similar to the behavior shown in
\cref{fig:pwr-keff}. With probability tables turned on, however, the situation
is quite different. At zero days, $k_\text{eff}$ calculated by OpenMC is 20 pcm
lower than the value calculated by Serpent. As the burnup increases, there is an
increasing positive bias between OpenMC and Serpent. To understand the cause of
this difference, it is instructive to look at differences in isotopic
compositions.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/sfr_keff.pdf}
  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
  between OpenMC and Serpent for the SFR assembly model. While not shown, the one
  standard deviation uncertainty in the difference is about 3.5 pcm.}
  \label{fig:sfr-keff}
\end{figure}

@@

With probability tables turned on, the differences between OpenMC and Serpent in
fission product concentrations are mostly the same as when probability tables
were turned off, as shown in \cref{fig:sfr-fp-nop}. Because the difference in
$^{238}$U and $^{239}$Pu concentrations was less than 0.2\% in
\cref{fig:sfr-actinides}, it makes sense that the use of probability tables
would not have a large effect on the fission product concentrations. Again, most
fission product concentrations agree to within 1\%, with $^{109}$Ag and
$^{85}$Kr being the notable outliers.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average_nop.pdf}
    \caption{Full depletion chain}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average_nop.pdf}
    \caption{CASL depletion chain}
  \end{subfigure}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables off.}
  \label{fig:sfr-fp-nop}
\end{figure}

@@

A new depletion solver has been implemented in the OpenMC Monte Carlo particle
transport code. The depletion solver is implemented primarily in Python 3 and
interfaces with OpenMC's transport solver through Python bindings to an
underlying C++ API, which enables data transfers to happen in memory rather than
through filesystem I/O. The design of the depletion solver relies on a set of
abstract classes that provide modularity and separation between different
components, including the time integration methods, the matrix exponential
solver, and the transport operator. In principle, this modularity would allow
the depletion system to be used with another transport code. Rather than relying
on a single time integration method, OpenMC implements several time integration
schemes including the predictor method, standard predictor--corrector methods,
stochastic implicit Euler integrators, and several novel
methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
OpenMC relies on the incomplete partial fraction form of CRAM by default.

@@

When comparing results on the SFR assembly model, larger differences in
$k_\text{eff}$ and isotopic concentrations were observed when unresolved
resonance region probability tables were turned on. These differences were found
to be related to how the temperature dependence of cross sections is treated in
the unresolved resonance region. In OpenMC, when the temperature of a material
is inbetween two temperatures at which probability tables are tabulated,
statistical interpolation between the tables is performed. Serpent, on the other
hand, uses probability tables at the lower tabulated temperature. For the SFR
model, this results in lower parasitic neutron absorption in $^{238}$U, lower
production of $^{239}$Pu, and ultimately a lower $k_\text{eff}$. After depleting
for 360 days, there was a 100 pcm difference between OpenMC and Serpent due to
this effect. Although this is not a huge difference, users who are simulating
problems with a non-uniform temperature distribution should be aware of this
behavior.

@@

In the present work, we have only performed a cross-code comparison between
OpenMC and Serpent. While the results are certainly encouraging, we would
caution readers that no comparisons against experimental results have been
performed to date and would be needed to ensure the fitness for a particular
application. We also highlight the fact that the present comparisons are focused
solely on fission reactor applications. Although the \texttt{openmc.deplete}
module may eventually be of interest to other application areas, further changes
and enhancements would be needed. For example, for use in fusion applications, a
wider variety of transmutation reactions would need to be accounted for, and the
ability to handle transmutation from a fixed, pulsed neutron source would be
required.
